---
title: "week3 Project"
author: "Nadia Gontova"
date: "2023-11-30"
output:
  pdf_document: default
  html_document: default
---


### DATA

The data used is an NYPD dataset covering a vareity of information on every recorded shooting from 2006 to 2021. The data includes information about the location, victim, and shooter. This is a very large dataset with almost 3000 entires however there is some missing information. Information on the perpetrator is often missing from entries in the dataset. 

```{r get_shooting_data}

if (!require("dplyr")) install.packages("dplyr")
if (!require("ggplot2")) install.packages("ggplot2")
library(dplyr)
library(ggplot2)

# URL
url_in <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"

# Read the data 
data <- read.csv(url_in) 

# removed
columns_to_remove <- c("Latitude", "Longitude", "X_COORD_CD", "Y_COORD_CD", "Lon_Lat", "LOC_OF_OCCUR_DESC", "LOCATION_DESC", "LOC_CLASSFCTN_DESC", "INCIDENT_KEY", "STATISTICAL_MURDER_FLAG", "JURISDICTION_CODE")

#data <- data[, !(names(data) %in% columns_to_remove)]
#data <- data[data$PERP_AGE_GROUP != "UNKNOWN", ]


# Convert OCCUR_DATE to date object
data$OCCUR_DATE <- as.Date(data$OCCUR_DATE, format = "%m/%d/%Y")

# Convert OCCUR_TIME to time data type
data$OCCUR_TIME <- as.POSIXct(data$OCCUR_TIME, format = "%H:%M", tz = "UTC")


head(data)

summary(data)
```


## Analysis 

```{r shootings_analysis}
# Group by BORO and calculate # of shootings in each place
shootings_per_boro <- data %>%
  group_by(BORO) %>%
  summarise(NumberOfShootings = n())

print(shootings_per_boro)

# Group data by OCCUR_DATE 
shootings_over_time <- data %>%
  group_by(OCCUR_DATE) %>%
  summarise(NumberOfShootings = n())

print(shootings_over_time)

```

### Plots



```{r shootings_plots}

# Bar plot
bar_plot <- ggplot(shootings_per_boro, aes(x = BORO, y = NumberOfShootings)) +
  geom_bar(stat = "identity", fill = "hotpink") +
  labs(title = "Shootings by Borough",
       x = "Borough",
       y = "# of Shootings")

print(bar_plot)

# line plot
line_plot <- ggplot(shootings_over_time, aes(x = OCCUR_DATE, y = NumberOfShootings)) +
  geom_line(color = "pink") +
  labs(title = "# of Shootings Over Time",
       x = "Date",
       y = "# of Shootings") +
  theme_classic()

print(line_plot)

```


```{r model}

shootings_over_time$Time <- 1:nrow(shootings_over_time)

# linear regression model
linear_model <- lm(NumberOfShootings ~ Time, data = shootings_over_time)

# Prediction
futuretime <- seq(max(shootings_over_time$Time) + 1, length.out = 365, by = 1)

futuredata <- data.frame(Time=futuretime)

predicted_values <- predict(linear_model, newdata=futuredata)

# Plot
plot(shootings_over_time$OCCUR_DATE, shootings_over_time$NumberOfShootings, 
     type = "l", col = "pink", xlab = "Date", ylab = "# of Shootings",
     main = "Linear Regression")
lines(shootings_over_time$OCCUR_DATE, fitted(linear_model), col = "red", lty = 2)
lines(seq(max(shootings_over_time$OCCUR_DATE) + 1, length.out = 365, by = 1), predicted_values, col = "black", lty = 2)
legend("topright", legend = c("Observed", "Fitted", "Predicted"), col = c("pink", "red", "black"), lty = c(1, 2, 2))


brooklyn_data <- data %>%
  filter(BORO == "BROOKLYN")

# Group data by OCCUR_DATE for Brooklyn
shootings_over_time_brooklyn <- brooklyn_data %>%
  group_by(OCCUR_DATE) %>%
  summarise(NumberOfShootings = n())

# Linear regression model for Brooklyn
linear_model_brooklyn <- lm(NumberOfShootings ~ as.numeric(OCCUR_DATE - min(shootings_over_time_brooklyn$OCCUR_DATE)), 
                            data = shootings_over_time_brooklyn)

# Prediction for Brooklyn
futuretime_brooklyn <- seq(max(shootings_over_time_brooklyn$OCCUR_DATE) + 1, length.out = 365, by = 1)
futuredata_brooklyn <- data.frame(OCCUR_DATE = futuretime_brooklyn)
predicted_values_brooklyn <- predict(linear_model_brooklyn, newdata = futuredata_brooklyn)

# Plot for Brooklyn
plot(shootings_over_time_brooklyn$OCCUR_DATE, shootings_over_time_brooklyn$NumberOfShootings, 
     type = "l", col = "pink", xlab = "Date", ylab = "# of Shootings",
     main = "Linear Regression for Brooklyn")
lines(shootings_over_time_brooklyn$OCCUR_DATE, fitted(linear_model_brooklyn), col = "red", lty = 2)
lines(futuretime_brooklyn, predicted_values_brooklyn, col = "black", lty = 2)
legend("topright", legend = c("Observed", "Fitted", "Predicted"), col = c("pink", "red", "black"), lty = c(1, 2, 2))

bronx_data <- data %>%
  filter(BORO == "BRONX")

# Group data by OCCUR_DATE for the Bronx
shootings_over_time_bronx <- bronx_data %>%
  group_by(OCCUR_DATE) %>%
  summarise(NumberOfShootings = n())

# Linear regression model for the Bronx
linear_model_bronx <- lm(NumberOfShootings ~ as.numeric(OCCUR_DATE - min(shootings_over_time_bronx$OCCUR_DATE)), 
                         data = shootings_over_time_bronx)

# Prediction for the Bronx
futuretime_bronx <- seq(max(shootings_over_time_bronx$OCCUR_DATE) + 1, length.out = 365, by = 1)
futuredata_bronx <- data.frame(OCCUR_DATE = futuretime_bronx)
predicted_values_bronx <- predict(linear_model_bronx, newdata = futuredata_bronx)

# Plot for the Bronx
plot(shootings_over_time_bronx$OCCUR_DATE, shootings_over_time_bronx$NumberOfShootings, 
     type = "l", col = "pink", xlab = "Date", ylab = "# of Shootings",
     main = "Linear Regression for Bronx")
lines(shootings_over_time_bronx$OCCUR_DATE, fitted(linear_model_bronx), col = "red", lty = 2)
lines(futuretime_bronx, predicted_values_bronx, col = "black", lty = 2)
legend("topright", legend = c("Observed", "Fitted", "Predicted"), col = c("pink", "red", "black"), lty = c(1, 2, 2))
```

#Conclusion 

In my analysis I found that Brooklyn and Bronx had the most shootings over the other regions. Staten Island had the least shootings from the five locations. In the graph of shootings over time I did not see any trend so I decided to preform a linear regression on the data. The linear model did not show an increase or decrease in shootings over time and based on this pattern the predicted pattern does not continue to increase or decrease. This aligns with what I visually saw when looking at the data. It is possible that the shootings data follows a different kind of trend and is therefore not well fit to a linear model. 

#biases

One source of bias could be from bias in the data collection. There is no way for me to verify how accuratly the data is collected and if certain entires might be missing. There might be a bais in how the regions are grouped and a small area of crime might not be representative of an entire region. Frthermore shootings in one region could possibly skew the data in a neighboring region due to how the borders are determined. These typed of biases can be reduced by looking into how the data is collected and how accurate the dataset is. 

There are also personal biases as an analyist that might skew the data. For example, knowing certain regions might have higher crime could impact the analysis someone chooses. To reduce this bias I tried to stick to analyzing the dataset as a whole and sticking to know statisical methods from class. 